---
title: A05
author: R package build
date: '2021-06-15'
slug: a05
categories: []
tags: []
---
## Import Libraries
```{r packages, echo = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(ggtext)
library(knitr)
library(kableExtra)
set.seed(1234)
options(dplyr.print_min = 10, dplyr.print_max = 6)
```

## Data: Class Sizes
```{r message = FALSE}
ClassSize <- read_csv("Class Size District Level Summary.csv", na = c("n/a", "", "NA")) %>%
   rename(
    NumberOfStudents = "NUMBER OF STUDENTS / SEATS FILLED",
    AvgClassSize = "AVERAGE CLASS SIZE",
    NumberSections = "NUMBER OF SECTIONS"
   )

ClassSize
```


- Number of observations: `r nrow(ClassSize)`
- Number of variables: `r ncol(ClassSize)`

---

## Goal: Predict average class size based on total number of students in school

$$\widehat{Average Class Size}_{i} = \beta_0 + \beta_1 \times Total Students In School_{i}$$


```{r students-classsize-plot, echo=FALSE, warning=FALSE}
ggplot(ClassSize, aes(x = NumberOfStudents, y = AvgClassSize)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "#8E2C90") + 
  labs(
    title = "Average class size versus number of students in school",
    subtitle = "City of New York 2010-2011",
    x = "Number of students in school",
    y = "Average class size"
  )
```

## Fit model & estimate parameters
```{r fit-model}
linear_reg() %>%
  set_engine("lm") %>%
  fit(AvgClassSize ~ NumberOfStudents, data = ClassSize) %>%
  tidy()
```

## A closer look at model output

```{r ref.label="fit-model", echo=FALSE}
```

.large[
$$\widehat{AverageClassSize}_{i} = 21.101 + 0.00146 \times TotalStudentsInSchool_{i}$$
]

---

## Slope and intercept

.large[
$$\widehat{AverageClassSize}_{i} = 21.101 + 0.00146 \times TotalStudentsInSchool_{i}$$
]

--

- **Slope:** For each additional student added to the total students in the school, the average class size is expected to increase by 0.00146 students.

--
- **Intercept:** Schools that have zero students are expected to have an average of 21.101 students in a class. This doesn't make sense, as if there were no students in a school, there would be no students to host in a class.

---

## Least squares regression

- The regression line minimizes the sum of squared residuals.

--
- If $e_i = y_i - \hat{y}_i$, then, the regression line minimizes 
$\sum_{i = 1}^n e_i^2$.

---

## Visualizing residuals

```{r vis-res-1, echo=FALSE, out.width="70%"}
cs_ss_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(AvgClassSize ~ NumberOfStudents, data = ClassSize)
cs_ss_fit_tidy <- tidy(cs_ss_fit$fit) 
cs_ss_fit_aug  <- augment(cs_ss_fit$fit) %>%
  mutate(res_cat = ifelse(.resid > 0, TRUE, FALSE))
p <- ggplot(data = cs_ss_fit_aug, 
            aes(x = NumberOfStudents, y = AvgClassSize)) +
  geom_point(alpha = 0.2) + 
  labs(
    title = "Average class size versus number of students in school",
    subtitle = "Just the data",
    x = "Number of students in school",
    y = "Average class size"
    ) +
  coord_cartesian(xlim = c(0, 15000), ylim = c(0, 50)) +
  theme(plot.subtitle = element_text(colour = "#E48957", face = "bold", size = rel(1.5)))
p
```

## Visualizing residuals (cont.)

```{r vis-res-2, echo=FALSE, out.width="70%"}
p <- p + 
  geom_smooth(method = "lm", color = "#8E2C90", se = FALSE) +
  geom_point(mapping = aes(y = .fitted), color = "#E48957") +
  labs(subtitle = "Data + least squares line")
p
```

## Visualizing residuals (cont.)

```{r vis-res-3, echo = FALSE, out.width="70%"}
p + 
  geom_segment(mapping = aes(xend = NumberOfStudents, yend = .fitted), color = "#E48957", alpha = 0.4) +
  labs(subtitle = "Data + least squares line + residuals")
```

## Properties of least squares regression

- The regression line goes through the center of mass point, the coordinates corresponding to average $x$ and average $y$, $(\bar{x}, \bar{y})$:  

$$\bar{y} = b_0 + b_1 \bar{x} ~ \rightarrow ~ b_0 = \bar{y} - b_1 \bar{x}$$

--
- The slope has the same sign as the correlation coefficient: $b_1 = r \frac{s_y}{s_x}$

--
- The sum of the residuals is zero: $\sum_{i = 1}^n e_i = 0$

--
- The residuals and $x$ values are uncorrelated

---

## Categorical predictor with 2 levels

## Relationship between average class size and district

.pull-left-narrow[
.small[
```{r echo=FALSE}
ClassSize %>% 
  select(CSD, AvgClassSize) %>%
  print(n = 20)
```

```{r cs-ns-fit}
linear_reg() %>%
  set_engine("lm") %>%
  fit(AvgClassSize ~ factor(CSD), data = ClassSize) %>%
  tidy()
```
## Predictor for district 2

$$\widehat{AvgClassSize} = 20.3 + 1.53~(CSD2)$$

- **Slope:** Classes in district 2 are expected, on average, to have 1.53 students more than classes in the baseline district 1.
    
- **Intercept:** Class sizes in district 1 are expected, on average, to have 20.3 students.

---
- When the categorical explanatory variable has many levels, they're encoded to **dummy variables**
- Each coefficient describes the expected difference between class sizes in that particular district compared to the baseline level

## Categorical predictor with 3+ levels

.pull-left-wide[
```{r echo=FALSE}
dummy_df <- ClassSize %>% 
  select(CSD) %>% 
  group_by(CSD) %>% 
  sample_n(1) %>%
  mutate(
    '1'   = as.integer(ifelse(CSD == "1", 1L, 0)),
    '2'    = as.integer(ifelse(CSD == "2", 1L, 0)),
    '3'    = as.integer(ifelse(CSD == "3", 1L, 0)),
    '4'    = as.integer(ifelse(CSD == "4", 1L, 0)),
    '5'    = as.integer(ifelse(CSD == "5", 1L, 0)),
    '6'    = as.integer(ifelse(CSD == "6", 1L, 0)),
    '7'    = as.integer(ifelse(CSD == "7", 1L, 0)),
    '8'    = as.integer(ifelse(CSD == "8", 1L, 0)),
    '9'    = as.integer(ifelse(CSD == "9", 1L, 0)),
    '10'    = as.integer(ifelse(CSD == "10", 1L, 0)),
    '11'    = as.integer(ifelse(CSD == "11", 1L, 0)),
    '12'    = as.integer(ifelse(CSD == "12", 1L, 0)),
    '13'    = as.integer(ifelse(CSD == "13", 1L, 0)),
    '14'    = as.integer(ifelse(CSD == "14", 1L, 0)),
    '15'    = as.integer(ifelse(CSD == "15", 1L, 0)),
    '16'    = as.integer(ifelse(CSD == "16", 1L, 0)),
    '17'    = as.integer(ifelse(CSD == "17", 1L, 0)),
    '18'    = as.integer(ifelse(CSD == "18", 1L, 0)),
    '19'    = as.integer(ifelse(CSD == "19", 1L, 0)),
    '20'    = as.integer(ifelse(CSD == "20", 1L, 0)),
    '21'    = as.integer(ifelse(CSD == "21", 1L, 0)),
    '22'    = as.integer(ifelse(CSD == "22", 1L, 0)),
    '23'    = as.integer(ifelse(CSD == "23", 1L, 0)),
    '24'    = as.integer(ifelse(CSD == "24", 1L, 0)),
    '25'    = as.integer(ifelse(CSD == "25", 1L, 0)),
    '26'    = as.integer(ifelse(CSD == "26", 1L, 0)),
    '27'    = as.integer(ifelse(CSD == "27", 1L, 0)),
    '28'    = as.integer(ifelse(CSD == "28", 1L, 0)),
    '29'    = as.integer(ifelse(CSD == "29", 1L, 0)),
    '30'    = as.integer(ifelse(CSD == "30", 1L, 0)),
    '31'    = as.integer(ifelse(CSD == "31", 1L, 0)),
    '32'    = as.integer(ifelse(CSD == "32", 1L, 0))
    
  )
dummy_df %>%
  kable(align = "lcccccc") %>%
  kable_styling() %>%
  column_spec(2, width = "10em", background = spec_color(dummy_df$'1'[1:32], end = 0.8), color = "white") %>%
  column_spec(3, width = "10em", background = spec_color(dummy_df$'2'[1:32], end = 0.8), color = "white") %>%
  column_spec(4, width = "10em", background = spec_color(dummy_df$'3'[1:32], end = 0.8), color = "white") %>%
  column_spec(5, width = "10em", background = spec_color(dummy_df$'4'[1:32], end = 0.8), color = "white") %>%
  column_spec(6, width = "10em", background = spec_color(dummy_df$'5'[1:32], end = 0.8), color = "white") %>%
  column_spec(7, width = "10em", background = spec_color(dummy_df$'6'[1:32], end = 0.8), color = "white") %>%
  column_spec(8, width = "10em", background = spec_color(dummy_df$'7'[1:32], end = 0.8), color = "white") %>%
  column_spec(9, width = "10em", background = spec_color(dummy_df$'8'[1:32], end = 0.8), color = "white") %>%
  column_spec(10, width = "10em", background = spec_color(dummy_df$'9'[1:32], end = 0.8), color = "white") %>%
  column_spec(11, width = "10em", background = spec_color(dummy_df$'10'[1:32], end = 0.8), color = "white") %>%
  column_spec(12, width = "10em", background = spec_color(dummy_df$'11'[1:32], end = 0.8), color = "white") %>%
  column_spec(13, width = "10em", background = spec_color(dummy_df$'12'[1:32], end = 0.8), color = "white") %>%
  column_spec(14, width = "10em", background = spec_color(dummy_df$'13'[1:32], end = 0.8), color = "white") %>%
  column_spec(15, width = "10em", background = spec_color(dummy_df$'14'[1:32], end = 0.8), color = "white") %>%
  column_spec(16, width = "10em", background = spec_color(dummy_df$'15'[1:32], end = 0.8), color = "white") %>%
  column_spec(17, width = "10em", background = spec_color(dummy_df$'16'[1:32], end = 0.8), color = "white") %>%
  column_spec(18, width = "10em", background = spec_color(dummy_df$'17'[1:32], end = 0.8), color = "white") %>%
  column_spec(19, width = "10em", background = spec_color(dummy_df$'18'[1:32], end = 0.8), color = "white") %>%
  column_spec(20, width = "10em", background = spec_color(dummy_df$'19'[1:32], end = 0.8), color = "white") %>%
  column_spec(21, width = "10em", background = spec_color(dummy_df$'20'[1:32], end = 0.8), color = "white") %>%
  column_spec(22, width = "10em", background = spec_color(dummy_df$'21'[1:32], end = 0.8), color = "white") %>%
  column_spec(23, width = "10em", background = spec_color(dummy_df$'22'[1:32], end = 0.8), color = "white") %>%
  column_spec(24, width = "10em", background = spec_color(dummy_df$'23'[1:32], end = 0.8), color = "white") %>%
  column_spec(25, width = "10em", background = spec_color(dummy_df$'24'[1:32], end = 0.8), color = "white") %>%
  column_spec(26, width = "10em", background = spec_color(dummy_df$'25'[1:32], end = 0.8), color = "white") %>%
  column_spec(27, width = "10em", background = spec_color(dummy_df$'26'[1:32], end = 0.8), color = "white") %>%
  column_spec(28, width = "10em", background = spec_color(dummy_df$'27'[1:32], end = 0.8), color = "white") %>%
  column_spec(29, width = "10em", background = spec_color(dummy_df$'28'[1:32], end = 0.8), color = "white") %>%
  column_spec(30, width = "10em", background = spec_color(dummy_df$'29'[1:32], end = 0.8), color = "white") %>%
  column_spec(31, width = "10em", background = spec_color(dummy_df$'30'[1:32], end = 0.8), color = "white") %>%
  column_spec(32, width = "10em", background = spec_color(dummy_df$'31'[1:32], end = 0.8), color = "white")
```

]
.pull-right-narrow[
.small[
```{r echo=FALSE}
ClassSize %>% 
  select(CSD, AvgClassSize,NumberOfStudents) %>%
  print(n = 20)
```

]
]

---

## Relationship between class size and district

### Rounding down for all decimal numbers as it's not possible to have partial students

.small[
```{r ref.label="cs-csd", echo=FALSE}
```

- **District 1** class sizes are expected, on average, to have **20** students.
- **District 2** class sizes are expected, on average, to have **1 more** student than *District 1*.
- **District 3** class sizes are expected, on average, to have **2 more** students than *District 1*.
- **District 4** class sizes are expected, on average, to have **no more** students than *District 1*.
- **District 5** class sizes are expected, on average, to have **no more** students than *District 1*.
- **District 6** class sizes are expected, on average, to have **2 more** students than *District 1*.
- **District 7** class sizes are expected, on average, to have **no more** students than *District 1*.
- **District 8** class sizes are expected, on average, to have **1 more** student than *District 1*.
- **District 9** class sizes are expected, on average, to have **1 more** student than *District 1*.
- **District 10** class sizes are expected, on average, to have **1 more** student than *District 1*.
- **District 11** class sizes are expected, on average, to have **2 more** students than *District 1*.
- **District 12** class sizes are expected, on average, to have **no more** students than *District 1*.
- **District 13** class sizes are expected, on average, to have **no more** students than *District 1*.
- **District 14** class sizes are expected, on average, to have **1 more** student than *District 1*.
- **District 15** class sizes are expected, on average, to have **1 more** student than *District 1*.
- **District 16** class sizes are expected, on average, to have **no more** students than *District 1*.
- **District 17** class sizes are expected, on average, to have **2 more** students than *District 1*.
- **District 18** class sizes are expected, on average, to have **2 more** students than *District 1*.
- **District 19** class sizes are expected, on average, to have **no more** students than *District 1*.
- **District 20** class sizes are expected, on average, to have **3 more** students than *District 1*.
- **District 21** class sizes are expected, on average, to have **3 more** students than *District 1*.
- **District 22** class sizes are expected, on average, to have **4 more** students than *District 1*.
- **District 23** class sizes are expected, on average, to have **no more** students than *District 1*.
- **District 24** class sizes are expected, on average, to have **2 more** students than *District 1*.
- **District 25** class sizes are expected, on average, to have **3 more** students than *District 1*.
- **District 26** class sizes are expected, on average, to have **4 more** students than *District 1*.
- **District 27** class sizes are expected, on average, to have **2 more** students than *District 1*.
- **District 28** class sizes are expected, on average, to have **3 more** students than *District 1*.
- **District 29** class sizes are expected, on average, to have **4 more** students than *District 1*.
- **District 30** class sizes are expected, on average, to have **2 more** students than *District 1*.
- **District 31** class sizes are expected, on average, to have **3 more** students than *District 1*.
- **District 32** class sizes are expected, on average, to have **1 more** student than *District 1*.
]
