---
title: "A03"
author: "R package build"
date: '2021-06-03'
slug: a03
categories: []
tags: []
---

```{r message=FALSE, echo=FALSE}
library(tidyverse)
lego_sales <- read_csv("lego_sales.csv")
```

We are exploring a data set that summarizes lego sales. It provides an information on who the customer was, their age, what theme of lego they bought, price, and several other variables. This allows us to analyze the data for trends and to gain insight into commonly purchased lego sets. 

## 1. What are the three most common first names of customers?
```{r message=FALSE, echo=FALSE}
lego_sales %>%
  count(first_name, sort = TRUE)
```
The three most common first names are Jackson, Jacob and Joseph.

## 2. What are the three most common themes of lego sets purchased?
```{r message=FALSE, echo=FALSE}
lego_sales %>%
  count(theme, sort = TRUE)
```
The three most common themes of lego sets purchased are Star Wars, Nexo Knights, and a tie for Gear and Mixels.

## 3. Among the most common theme of lego sets purchased, what is the most common subtheme?
```{r message=FALSE, echo=FALSE}
lego_sales %>%
  drop_na() %>%
  count(theme,subtheme, sort=TRUE) %>%
  filter(theme == 'Star Wars')
```
From the results of the question prior, we know that Star Wars was the most common theme of lego sets purchased. Of that theme, the most common subtheme is The Force Awakens.

## 4. Create a new variable called "age_group" and group the ages into categories
```{r message=FALSE, echo=FALSE}
age_group <- lego_sales %>%
  mutate(AgeGroup = case_when(
    age < 19 ~ "18 and under",
    between(age, 19, 25) ~ "19 - 25",
    between(age, 26, 35) ~ "26 - 35",
    between(age, 36, 50) ~ "36 - 50",
    age > 50 ~ "51 and over"
  ))

age_group

```
## 5. What is the probability a randomly selected customer?

### a. Is in the 19 - 25 age group?
```{r message=FALSE, echo=FALSE}
InAgeGroup <- age_group %>%
  count(AgeGroup) %>%
  mutate(prob = n/sum(n)) %>%
  filter(AgeGroup == '19 - 25')

InAgeGroup
```

There is a 20.8% chance that a randomly selected customer will be in the 19 - 25 age group.

### b. Is in the 19 - 25 age group and purchased a Duplo theme set?
```{r message=FALSE, echo=FALSE}
DuploTheme <- age_group %>%
  filter(AgeGroup == '19 - 25') %>%
  count(theme) %>%
  mutate(prob = n/sum(n)) %>%
  filter(theme == 'Duplo')

df <- cbind(InAgeGroup, DuploTheme)
df 


p <- df[1,3]*df[1,6]
p
```
The probability of a customer being in the 19 - 25 years age range AND purchased a Duplo themed set is 1.45%.

### c. Is in the 19 - 25 age group given they purchased a Duplo theme set?

```{r message=FALSE, echo=FALSE}
DuploTheme2 <- age_group %>%
  filter(theme == 'Duplo') %>%
  count(AgeGroup) %>%
  mutate(prob = n/sum(n)) %>%
  filter(AgeGroup == '19 - 25')

DuploTheme2
```
Given that the customer purchased a Duplo theme lego set, there is a 25.7% chance that they are in the 19 - 25 age range.

## 6. Which age group has purchased the largest number of lego sets? How many did they purchase?

```{r message=FALSE, echo=FALSE}
MostLego <- age_group %>%
  group_by(AgeGroup) %>%
  summarise(quantity = sum(quantity))

MostLego
```

The age group that purchased the largest number of lego sets is the 36 - 50 years, with a total of 313 sets bought.

## 7. Which age group has spent the most money on legos? How much did they spend?
```{r message=FALSE, echo=FALSE}
MostMoney <- age_group %>%
  group_by(AgeGroup) %>%
  summarise(us_price = sum(us_price))

MostMoney
```

The age group that has spent the most money on lego is also the age 36 - 50 group. They have spent a total of $6641.84

## 8. Custom question: What is the average sale price by age group?

```{r message=FALSE, echo=FALSE}
AverageSale <- age_group %>%
  group_by(AgeGroup) %>%
  summarise(us_price = mean(us_price))

ggplot(AverageSale, aes(x=AgeGroup, y=us_price)) + geom_col() +labs(x = "Age Group", y = "Average Sale Price", title= "Average Sale Price by Age Group")

```
This visualization shows that not only did the 36 - 50 year olds spend the most money total on lego, they also had the highest average spend per transaction. It can also be noted that the 51 and over group had the second highest average spend per transaction, but was not the second highest spender overall.

## 9. Added element to the data visualization plot

The elements that have been changed in this plot are:

* The plot bars are color filled based on age group
* The background color is now grey with white lines
* The legend has been moved to the bottom

```{r message=FALSE, echo=FALSE}
AverageSale <- age_group %>%
  group_by(AgeGroup) %>%
  summarise(us_price = mean(us_price))

ggplot(AverageSale, aes(x=AgeGroup, y=us_price, fill=AgeGroup)) + 
  geom_col() +
  labs(x = "Age Group", y = "Average Sale Price", title= "Average Sale Price by Age Group") +
  theme_grey() +
  theme(legend.position = "bottom")
 

```
## Discussion

This assignment was a great opportunity to practice newly learned skills in R Studio. It tested our abilities in extracting relevant data, and grouping variables as needed to answer the above questions. There were a few challenges encountered when attempting to extract specific data to calculate the probability questions especially, but overall, the assignment was the right amount of challenging to test our skills.

The analysis of the data set provides insight into the most popular lego sets, common ages of those purchasing the lego set, and how much money is spent overall on lego in the year 2018. 
