---
title: A03
author: R package build
date: '2021-06-03'
slug: a03
categories: []
tags: []
---

```{r}
library(tidyverse)
lego_sales <- read_csv("lego_sales.csv")
```

## 1. What are the three most common first names of customers?
```{r}
lego_sales %>%
  count(first_name, sort = TRUE)
```
The three most common first names are Jackson, Jacob and Joseph.

## 2. What are the three most common themes of lego sets purchased?
```{r}
lego_sales %>%
  count(theme, sort = TRUE)
```
The three most common themes of lego sets purchased are Star Wars, Nexo Knights, and a tie for Gear and Mixels.

## 3. Among the most common theme of lego sets purchased, what is the most common subtheme?
```{r}
lego_sales %>%
  count(theme, subtheme, sort=TRUE)
```

## 4. Create a new variable called "age_group" and group the ages into categories
```{r}
AgeGroup <- lego_sales %>%
  mutate(age_group = case_when(
    age < 19 ~ "18 and under",
    between(age, 19, 25) ~ "19 - 25",
    between(age, 26, 35) ~ "26 - 35",
    between(age, 36, 50) ~ "36 - 50",
    age > 50 ~ "51 and over"
  ))

AgeGroup
```
## 5. What is the probability a randomly selected customer?

### a. Is in the 19 - 25 age group?
```{r}

```

### b. Is in the 19 - 25 age group and purchased a Duplo theme set?
```{r}

```

### c. Is in the 19 - 25 age group given they purchased a Duplo theme set?

```{r}

```



