---
title: A05 v3
author: R package build
date: '2021-06-17'
slug: a05-v3
categories: []
tags: []
---
```{r message=FALSE, echo=FALSE}
library(tidyverse)
library(dplyr)
library(ggridges)

WeatherData <- read_csv("weather.csv")

WeatherData[WeatherData == 0] <- NA
``` 

## Dataset: US Weather

This dataset provides information on the weather throughout the United States in the year 2017. The data includes:

* Weather Station ID
* State
* Latitude
* Longitude
* Elevation
* Date
* Minimum, Maximum and Average Temperatures
* Average Daily Wind Speed
* Direction of Fastest 5-sec Wind
* Fastest 5-sec Wind Speed
* Snowfall
* Snow Depth
* Precipitation

## Initial Analysis Questions
Overarching question: Is there a single state with that experienced all of the most extreme weather events? Ie. Most snowfall, most precipitation, highest wind speeds, etc.

1. What is the coldest state?

2. Which state has the most snowfall?

3. Which state has the most precipitation?

5. Are there any correlation in the weather data?



## Dataset Exploration

### Data Entries by State

We will first examine the quantities of data collected by state to investigate whether all states have similar amounts of data to work from, or if some states won't have enough data to draw conclusions with. 

```{r message=FALSE, echo=FALSE}
StateCount <- WeatherData %>%
  group_by(state) %>%
  count(state)

ggplot(StateCount, aes(x=state, y=n, fill=state)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
 
```
Based on the above graph, we can see that there are significantly different amounts of data collected for each location. 

### Data Quality

Other data quality issues include a large number of "NA" data, shown in the table below. The drop_na() function will be used throughout the analysis to remove any data that doesn't contribute to the overall calculations that are executed.

```{r message=FALSE, echo=FALSE}
WeatherData %>%
  count("NA")
```
Lastly, this is less of a data quality issue but rather something that has been noted about the data. The dataset is described as weather data from the United States, and refers to the locations as "states", however, the data also appears to capture Canadian provinces and territories. For the purpose of this analysis, where possible, the variable "state" will be more correctly referred to as "location" so as not to limit the location under investigation to one country.

## Data Analysis

### Coldest State

#### Coldest Temperature Experienced in the US

To understand whether there is a single location that experiences the all of the most extreme weather events, we will first explore the coldest logged temperature.
```{r message=FALSE, echo=FALSE}
ColdestTemp <- WeatherData %>%
  drop_na(TMIN) %>%
  group_by(state) %>%
  summarize(MinTemp = min(TMIN)) %>%
  arrange(MinTemp) %>% 
  head(10)

ggplot(ColdestTemp, aes(y = state, x = MinTemp, fill = state)) + geom_col() + labs(title="Coldest Logged Temperature Event", x="Temperature (F)", y="Location (US State or Canadian Prov/Terr.")
```
Here we see that the state that experienced the coldest temperature is Alaska. In the beginning, this temperature appeared to be a significant outlier, but perhaps this temperature is possible within this state due to it's location. 


#### Coldest State on Average

Taking the above analysis a step further and see if the location that experienced the coldest logged temperature is also the coldest on average. This was explored by taking the average temperature overall of each state. This demonstrated that the extreme temperatures summarized above do not necessarily indicate coldest average temperatures overall. 

In the average temperature analysis below, we see that the Northwest Territories is the coldest location overall. It should be noted that only the top 10 locations that experienced the coldest temperatures are reported on the graph for clarity.
```{r message=FALSE, echo=FALSE}
ColdestAvg <- WeatherData %>%
  drop_na(TAVG) %>%
  group_by(state) %>%
  summarize(AverageTemperature = mean(TAVG)) %>%
  arrange((AverageTemperature)) %>%
  head(10)


ggplot(ColdestAvg, aes(x=state, y=AverageTemperature, fill = state)) + geom_col() + labs(title="Coldest Location on Average", x="Temperature (F)", y="Location (US State or Canadian Prov/Terr.")

```
#### A Closer Look at the Coldest Location (on average)

The following visualization summarizes all the average temperatures taken throughout the year (average of the highest and lowest temperature each day) with the use of a boxplot. This allows us to observe the temperatures experienced in this location in more detail.
```{r message=FALSE, echo=FALSE}
ColdestState <- WeatherData %>%
  filter(state == "NT")

ggplot(ColdestState, aes(x = state, y = TAVG)) + geom_boxplot() + labs(title="Northwest Territories Temperatures", x="Northwest Territories", y="Average Temperature (F)")
```


### Snowfall

The next weather variable to analyze is the snowfall. This analysis will follow a similar procedure in that it will determine which location experienced the most significant snow fall, then determine which location has the most snow on average to see if they are the same. We will compare these results to the results obtained above to see if the location that experienced the coldest temperatures also experiences the most snow.

#### Most Snowfall

Here we see that the location that experienced the most snow is New Jersey. This differs from the location that experienced the coldest temperature, so our overaching question has already been answered. We can conclude that no, there is not a single state that experience all of the extreme weather events, but rather, each weather variable varies by location.
```{r message=FALSE, echo=FALSE}
MaxSnow <- WeatherData %>%
  drop_na(SNOW) %>%
  group_by(state) %>%
  summarize(MaxSnowfall = max(SNOW)) %>%
  arrange(desc(MaxSnowfall)) %>%
  head(10)

MaxSnow

ggplot(MaxSnow, aes(x = state, y = MaxSnowfall, fill=state)) + geom_col() + labs(title="Highest Logged Snowfall Event", x="State", y="Snowfall (inch)")

```
#### Most Snowfall on Average

To determine whether locations experiencing peak snowfall events also have the highest snowfall on average, the average total snowfall by location has been summarized and graphed below.

In this scenario, the location that experienced the highest snowfall event also experienced the highest snowfall on average for the entire year.
```{r message=FALSE, echo=FALSE}
MostSnowAvg <- WeatherData %>%
  drop_na(SNOW) %>%
  group_by(state) %>%
  summarize(AverageSnowfall = mean(SNOW)) %>%
  arrange(desc(AverageSnowfall)) %>%
  head(10)

MostSnowAvg

ggplot(MostSnowAvg, aes(x = state, y = AverageSnowfall, fill = state)) + geom_col() + labs(title="Most Snowfall on Average", x="State", y="Snowfall (inch)")
```


#### A Closer Look at Highest Average Snowfall Location

A line graph is used below to map out the snowfall throughout the winter months in the location that has experienced both the most extreme snowfall, as well as the most snow on average, New Jersey.

From the below graph, it appears that while there aren't a significant number of snowfall events, the events that have occurred are large, which has resulted in the highest snowfall on average above.
```{r message=FALSE, echo=FALSE}
MostSnow <- WeatherData %>%
  drop_na(SNOW) %>%
  filter(state == "NJ")
 
MostSnow

ggplot(MostSnow, aes(x = date, y = SNOW)) + geom_line() + labs(title="New Jersey Snowfall", x="Date", y="Snowfall (inch)")
```
### Precipitation

We will now analyze the precipitation data to see which location has experienced the highest rainfall event. We will then investigate this in a slightly different way, by looking at the 3 locations that experienced the highest rainfall events

#### Most Precipitation
```{r message=FALSE, echo=FALSE}
MostPrecip <- WeatherData %>%
  drop_na(PRCP) %>%
  group_by(state) %>%
  summarize(MaxPrecipitation = max(PRCP)) %>%
  arrange(desc(MaxPrecipitation)) %>%
  head(10)

MostPrecip

ggplot(MostPrecip, aes(x = state, y = MaxPrecipitation, fill=state)) + geom_col() + labs(title="Highest Logged Precipitation Event", x="State", y="Precipitation (inch)")
```
#### Top 3 Highest Precipitation Events

Based on the above plot, we see that the locations experiencing the highest precipitation events are Texas (TX), Florida (FL), and Alaska (AK). For further analysis, a boxplot has been used to better understand the spread of data for these locations. Are these precipitation events outliers, or is it common for them to have these levels of precipitation throughout the year?

```{r message=FALSE, echo=FALSE}
TXData <- WeatherData %>%
  drop_na(PRCP) %>%
  filter(state == 'TX')

FLData <- WeatherData %>%
  drop_na(PRCP) %>%
  filter(state == 'FL')

AKData <- WeatherData %>%
  drop_na(PRCP) %>%
  filter(state == 'AK')

ggplot() +
  geom_boxplot(data = TXData, aes(y=PRCP, x=state, fill=state)) + 
  geom_boxplot(data = FLData, aes(y=PRCP, x=state, fill=state)) +
  geom_boxplot(data = AKData, aes(y=PRCP, x=state, fill=state)) +
  labs(title = "Top 3 Locations with Highest Precipitation Events", x="State", y="Precipitation (inch)")
```

From the box plot, we see that typically, the precipitation events are only a few inches, but each of 3 locations experiences fairly frequent high precipitation events throughout the year.

#### A Closer Look at Yearly Texas Precipitation

To further examine the precipitation in Texas over the year, a line plot has been used to graph all of the recorded precipitation events. We can see that precipitation is frequent throughout the year in Texas, with highs around 5 inches of precipitation, and the one outlier that has exceeded 20 inches.
```{r message=FALSE, echo=FALSE}
ggplot(TXData, aes(x=date, y=PRCP)) + geom_line() +
  labs(title = "Texas Precipitation Events", x="Date", y="Precipitation (inch)")
```

### Correlations

The last aspect of the data we will analyze is whether there are any correlations between the weather variables that are recorded. 

#### Correlation Analysis 1: Snow versus Precipitation

Here we seek to understand if there is any correlation between snowfall and precipitation events. Do locations that experience high precipitation also experience high snowfall events?

Based on the below graph, we can see that there is some correlation between locations that experience high precipitation events also experiencing high snowfall events. 

```{r message=FALSE, echo=FALSE}
ggplot(WeatherData, aes(y=SNOW, x=PRCP)) + geom_point() +
  labs(title = "Correlation Between Snow vs. Precipitation Events", x="Precipitation (inch)", y="Snow (inch)") +xlim(0, 5) + ylim(0,40)
```
#### Correlation Analysis 2: Average Temperature versus Precipitation

When comparing the average temperature to precipitation events, we can see there isn't a clear linear correlation between these two variables. Interestingly enough, there also isn't a normal distribution of the data. The data most resembles a double normal distribution. Perhaps with data from additional years, a more clear correlation would appear between higher temperatures and higher precipitation events.
```{r message=FALSE, echo=FALSE}
ggplot(WeatherData, aes(x=TAVG,y=PRCP)) + geom_point(alpha=0.5) +
  labs(title = "Correlation Between Average Temperature vs. Precipitation Events", x="Average Temperature (F)", y="Precipitation (inch)")
```

#### Correlation Analysis 3: Average Temperature versus Snow

The final correlation investigated is whether there is a relationship between the average temperature versus snow fall events. The expectation was that the locations with lower average temperatures would be linearly correlated with larger snow fall events. The below graph shows that this isn't the case. the data here is normally distributed rather than normally correlated, which is an unexpected finding.

```{r message=FALSE, echo=FALSE}
ggplot(WeatherData, aes(x=TAVG,y=SNOW)) + geom_point()
```

## Discussion
This analysis provided insight weather experienced within the United States and Canada. The data allowed us to explore extreme weather events such as coldest temperatures, most snow fall and most precipitation and determine whether there were any correlations between the weather variables that are collected throughout the year 2017. 

The overarching question was whether there was a single state that experienced all of the most extreme weather events such as coldest temperature, most snow AND most precipitation. The analysis shows that this is not the case, and that the most extreme weather results were experienced in different locations, as follows:

* Coldest Temperature = Alaska
* Largest Snowfall Event = New Jersey
* Largest Precipitation Event = Texas

The analysis then looked into whether there were any strong correlations between weather event variables. The correlations determined were as follows:

* Snow versus Precipitation Events = Linear Correlation
* Average Temperature versus Precipitation Events = Double Normal Distribution
* Average Temperature versus Snowfall Events = Normal Distrbution

While there was a significant amount of "NA" variables within the dataset, the analysis was still able to provide insight into the weather in various locations and begin analyzing correlations between variables. This analysis could benefit from additional years, or perhaps more complete data from this single year, but still offered valuable information about weather experienced in different parts of the world.




